Network Fusion through CDRP Model\\ Decomposition and Assembly

### 2020-01-10 paper/paper.tex
some fix

Table~\ref{tab:assemble-example-encoding} shows the encoding mentioned in the certificate algorithm for the intro example in Figure~\ref{fig:assemble-example-venn}.
If we only use the encoding from $d_1, d_2, d_3$, we find no two subclasses with different labels share the same encoding, and therefore certify the correctness of the feasible solution proposed before.
For the partial classification task for $A_3$, using two bits from $d_2, d_4$ already provides distinguishable encoding for $A_{31}$ and the remaining dummy classes.

% \linespread{1.1}
\begin{table}[h]
	\begin{tabular}{|c|c|cccc|}
	\cline{1-6}
	\multicolumn{2}{|c|}{} & $d_1$ & $d_2$ & $d_3$ & $d_4$ \\ \cline{1-6}
	\multirow{2}{*}{$A_1$} & $A_{11}$ & 1 & 0 & 0 & 0 \\ \cline{2-6}
	 & $A_{12}$ & 1 & 0 & 0 & 0 \\ \cline{1-6}
	\multirow{3}{*}{$A_2$} & $A_{21}$ & 1 & 1 & 0 & 0 \\ \cline{2-6}
	 & $A_{22}$ & 1 & 0 & 1 & 1 \\ \cline{2-6}
	 & $A_{23}$ & 0 & 0 & 1 & 0 \\ \cline{1-6}
	$A_3$ & $A_{31}$ & 0 & 1 & 0 & 1 \\ \cline{1-6}
	\end{tabular}
	\caption{Encoding for the Intro Example}
	\label{tab:assemble-example-encoding}
\end{table}
% \linespread{1.0}

@@

\begin{algorithm}
	\caption{The General Algorithm}
	\label{tab:heuristic-search}
	\begin{algorithmic}[1]
		\Require The inputs to the assembly problem;
		\Ensure $D$, a classifier set satisfying constraints, or $\emptyset$ if not feasible.
		\State $Q \gets \emptyset$; \Comment A priority queue, the last element $h$ is the key
		\Function{Expand}{$D$} \Comment $D$, the classifier set
			\For{$d \in \mathcal{D} - D$}
				\State Calculate the encoding $C$ using $D \cup \{d\}$;
				\State Calculate the entropy $h$ for $C$;
				\State $Q \gets Q \cup \{(D \cup \{d\}, C, h)\}$;
			\EndFor
		\EndFunction
		\Function{HeuristicSearch}{}
			\State $\Call{Expand}{\emptyset}$;
			\While{$Q \neq \emptyset$}
				\State $(D, C, h) \gets Pop(Q)$;
				\If{$e_0 == 0$}
					\State \Return $D$;
				\EndIf
				\If{|D| < k}
					\State $\Call{Expand}{D}$;
				\EndIf
			\EndWhile
			\State \Return $\emptyset$;
		\EndFunction
	\end{algorithmic}
\end{algorithm}

@@

Critical Data Routing Path is a method to find a path containing critical nodes, whose removal causes great preformance setback, from a existing trained model. We introduce CDRP to decomposite one existing pretrained model from selecting representing CDRPs in a given dataset.

\paragraph{CDRP extraction}

From the original model, we introduce a scalar control gate to every output channel of one layer, with nonnegative and sparse value. They will be multiplied to the layer’s output channelwise, resulting in the routing nodes. The nodes that are not suppressed, or have control gate weight greater zero, will be linked together to compose the routing paths. Then finding the critical data routing path can be reduced to optimize the weights of the control gates.

Critical data routing path is extracted through optimizaition of control gates values. We freeze the weights of the model, and for each sampe we follow the procedure of Distillation Guided Routing (algorithm \ref{alg:dgr}) \cite{wang2018interpret} to get the gate values. The gate values can be concatenated into one single vector as representation of that sample. As the original paper \cite{wang2018interpret} states, these gate values encode patterns inside the input images, e.g. tinca in horizontal positions or squat positions will have different encodings.

\begin{algorithm}
	\caption{Distillation Guided Routing, introduced in paper \cite{wang2018interpret}}
	\label{alg:dgr}
	\begin{algorithmic}[1]
		\Require Input image $\boldsymbol{x}$, pretrained network $f_\theta(\cdot)$, control gates $\boldsymbol{\Lambda}$ initialized with 1, balanced parameter $\gamma$. Max iterations $T$, SGD optimizer
		\Ensure identified CDRPs representation $\boldsymbol{v}$
		\State original prediction class $i \gets \arg\max f_\theta(\boldsymbol{x})$
		\For{$t \gets 1$ to $T$}
			\State compute loss $cur\_loss$: $$\mathcal{L}\left(f_\theta(\boldsymbol{x}), f_\theta(\boldsymbol{x}; \boldsymbol{\Lambda})\right) + \gamma \sum_k |\boldsymbol{\lambda}_k|_1$$
			\State compute control gates gradients $\boldsymbol{\Lambda}$: $$\frac{\partial Loss}{\partial \boldsymbol{\Lambda}} = \frac{\partial \mathcal{L}}{\partial \boldsymbol{\Lambda}} + \gamma \times \mathrm{sign}(\boldsymbol{\Lambda})$$
			\State update $\boldsymbol{\Lambda}$ by SGD optimzier and clip $\boldsymbol{\Lambda}$ to be nonnegative
			\State new prediction class $j \gets \arg\max f_\theta(\boldsymbol{x};\boldsymbol{\Lambda})$
			\If{$i=j$} \Comment keep the prediction same
				\If{$cur\_loss$ is minimun}
					\State $\boldsymbol{v} \gets \mathrm{concatenate}(\boldsymbol{\Lambda})$
				\EndIf
			\EndIf
		\EndFor		
	\end{algorithmic}
\end{algorithm}

\paragraph{CDRP clustering}

The fact that CDRP as encoding of samples directly leads to the idea that decompoisite one large model into several smaller models, each of which processing one pattern of input data. Once the gate vectors of all samples is obtained, we use k-means clusterng to cluster these vectors to gather samples sharing similar patterns together.

The visualization of clustering results is presented in figure \ref{fig:vgg-cluster-pca}.

\begin{figure}[!htp]
	\centering
	\subfigure[Result of VGG-16(C)] {
		\includegraphics[width=\linewidth]{fig/vgg16c-cluster.png}
        \label{fig:vgg16c-cluster}
	}
	\subfigure[Result of VGG-19(D)] {
		\includegraphics[width=\linewidth]{fig/vgg16d-cluster.png}
        \label{fig:vgg16d-cluster}
    }
    \caption{Visualization of clustering results. The axes are the first two components of of PCA results on all extracted gate vectors. The figure in the top is the result of VGG-16 (type C) trained on cifar-100 dataset, and the one below is the result of VGG-16 (type D) fine-tuned on cifar-100 initialized with pretrained weights on ImageNet. Different colors represent different class labels in cifar-100 dataset, and the cluster centers are denoted as blue squares.}
    \label{fig:vgg-cluster-pca}
\end{figure}

From the visualization result, it can be observed that the clustering can distinguish samples of difference class lables, i.e. samples in one cluster are not prone to have different class labels. Besides, gate vectors extracted on the same data by different models differs.

The results shows we can derive a good basis of binary classifers through these clusters. First, the good mapping from cluster to class label means we can use these binary classifiers to construct subclasses mentioned in the beginning of this section. The fact that gate vectors from different models varies will prevent the trival case that cluster one-to-one map to class label, leading to substantial reduction on number of model parameters.

\paragraph{Model pruning and fine tuning}

We use methods introduced in paper \cite{LiDongYue} to prune the model according to gate vector of cluster centers (algorithm \ref{alg:prune}).

\begin{algorithm}
	\caption{Model Pruning of clusters}
	\label{alg:prune}
	\begin{algorithmic}[1]
		\Require input images of one given cluster $TargetImages$, pretrained network $f_\theta(\cdot)$, $prune\_ratio$
		\Ensure Pruned neural network $f_{\theta'}(\cdot)$
		\State $Paths = \{\}$
		\For{each given training image $\boldsymbol{x}$ in $TargetImages$}
			\State $Path_{\boldsymbol{x}} = \mathrm{DistillationGuidedRouting}(\boldsymbol{x}, f_\theta(\cdot))$
			\State $Paths.add(Path_{\boldsymbol{x}})$
		\EndFor		
		\State compute cluster-wise critical data routing path $SumPath$ via merging $Paths$, resulting gate vector $\boldsymbol{\Lambda} = [\boldsymbol{\lambda}_k]$
		\For{$\boldsymbol{\lambda}_k$ in the $k$-th convolution layers}
			\State sort all $\lambda$s in $\boldsymbol{\lambda}_k$ decreasingly and discard the channels in the last within range of $prune\_ration$
		\EndFor
	\end{algorithmic}
\end{algorithm}

Afterwards we modify the network architecture of fully connected layers to form a binary classifer $g_{\theta'}(\cdot)$. The reason to do this includes:
\begin{itemize}
	\item It has less parameters than 100-class classifers.
	\item It is easier to train, i.e. easier to have higher accuarcy.
	\item Weights of convolution layers in the orginal model could be reused. We just need to finetune the model without running long sessions to train from scratch.
\end{itemize}

We finetune this model with samples having positive labels in the clusterwhich is used to prune the model, and all other samples having negative labels.

@@

\begin{table}[!htp]
	\begin{tabular}{|c|c|c|}
	\hline
	\multirow{2}{*}{Classifier} &\multicolumn{2}{|c|}{Possible Class Labels}\\
	\cline{2-3}
	& In (0) & Not in (1)\\
	\hline
	classifier-0,6 & 4 & 0,1,2,3,4,5\\
	classifier-1 & 1 & 0,2,3,4,5\\
	classifier-2 & 2 & 0,1,3,4,5\\
	classifier-3 & 0 & 1,2,3,4,5\\
	classifier-4 & 3 & 0,1,2,4,5\\
	classifier-7 & 3,4 & 0,1,2,3,4,5\\ 
	classifier-9 & 2,3,4 & 0,1,5\\ 
	classifier-8,13 & 0 & 1,2,3,4,5\\
	classifier-10,12 & 1 & 0,1,2,3,4,5\\
	classifier-11 & 3 & 0,1,2,3,4,5\\
	\hline

	\end{tabular}
	\caption{Corresponding Class Labels of Each Cluster}
	\label{tab:cluster_label}
\end{table}

@@

\begin{table}[!htp]
	\begin{tabular}{|c|c|c|c|c|}
	\hline
	& classifier-3 & classifier-4 & classifier-1 & classifier-2 \\
	\hline
	Accuracy & 96.0\% & 86.0\% & 91.5\% & 87.0\%\\
	\hline
	\end{tabular}
    \caption{Accuracy of Chosen Binary-Classifiers}
    \label{tab:classifer-accuracy}
\end{table}

@@

\begin{table}[!htp]
	\begin{tabular}{|c|c|c|c|c|c|c|}
	\hline
	& class-0 & class-1 & class-2 & class-3 & class-4 & Overall \\
	\hline
	Encoding &0 & 110 & 1110 &10 & 1111 &\\
	\hline
	\end{tabular}
    \caption{Encoding and Accuracy of Each Class Label}
    \label{tab:encoding-and-accuarcy}
\end{table}

>>>>

Figure~\ref{fig:assemble-example-encoding} shows the encoding mentioned in the certificate algorithm for the intro example in Figure~\ref{fig:assemble-example-venn}.
If we only use the encoding from $d_1, d_2, d_3$, we find no two subclasses with different labels share the same encoding, and therefore certify the correctness of the feasible solution proposed before.
For the partial classification task for $A_3$, using two bits from $d_2, d_4$ already provides distinguishable encoding for $A_{31}$ and the remaining dummy classes.

% \linespread{1.1}
\begin{figure}[h]
	\begin{tabular}{|c|c|cccc|}
	\cline{1-6}
	\multicolumn{2}{|c|}{} & $d_1$ & $d_2$ & $d_3$ & $d_4$ \\ \cline{1-6}
	\multirow{2}{*}{$A_1$} & $A_{11}$ & 1 & 0 & 0 & 0 \\ \cline{2-6}
	 & $A_{12}$ & 1 & 0 & 0 & 0 \\ \cline{1-6}
	\multirow{3}{*}{$A_2$} & $A_{21}$ & 1 & 1 & 0 & 0 \\ \cline{2-6}
	 & $A_{22}$ & 1 & 0 & 1 & 1 \\ \cline{2-6}
	 & $A_{23}$ & 0 & 0 & 1 & 0 \\ \cline{1-6}
	$A_3$ & $A_{31}$ & 0 & 1 & 0 & 1 \\ \cline{1-6}
	\end{tabular}
	\caption{Encoding for the Intro Example}
	\label{fig:assemble-example-encoding}
\end{figure}
% \linespread{1.0}

@@

\begin{algorithm}[h]
	\caption{The General Algorithm}
	\label{tab:heuristic-search}
	\begin{algorithmic}[1]
		\Require The inputs to the assembly problem;
		\Ensure $D$, a classifier set satisfying constraints, or $\emptyset$ if not feasible.
		\State $Q \gets \emptyset$; \Comment A priority queue, the last element $h$ is the key
		\Function{Expand}{$D$} \Comment $D$, the classifier set
			\For{$d \in \mathcal{D} - D$}
				\State Calculate the encoding $C$ using $D \cup \{d\}$;
				\State Calculate the entropy $h$ for $C$;
				\State $Q \gets Q \cup \{(D \cup \{d\}, C, h)\}$;
			\EndFor
		\EndFunction
		\Function{HeuristicSearch}{}
			\State $\Call{Expand}{\emptyset}$;
			\While{$Q \neq \emptyset$}
				\State $(D, C, h) \gets Pop(Q)$;
				\If{$e_0 == 0$}
					\State \Return $D$;
				\EndIf
				\If{|D| < k}
					\State $\Call{Expand}{D}$;
				\EndIf
			\EndWhile
			\State \Return $\emptyset$;
		\EndFunction
	\end{algorithmic}
\end{algorithm}

@@

@@

\begin{table}[h]
	\caption{Corresponding Class Labels of Each Cluster}
	\label{tab:cluster_label}
	\begin{tabular}{|c|c|c|}
	\hline
	\multirow{2}{*}{Classifier} &\multicolumn{2}{|c|}{Possible Class Labels}\\
	\cline{2-3}
	& In (0) & Not in (1)\\
	\hline
	classifier-0,6 & 4 & 0,1,2,3,4,5\\
	classifier-1 & 1 & 0,2,3,4,5\\
	classifier-2 & 2 & 0,1,3,4,5\\
	classifier-3 & 0 & 1,2,3,4,5\\
	classifier-4 & 3 & 0,1,2,4,5\\
	classifier-7 & 3,4 & 0,1,2,3,4,5\\ 
	classifier-9 & 2,3,4 & 0,1,5\\ 
	classifier-8,13 & 0 & 1,2,3,4,5\\
	classifier-10,12 & 1 & 0,1,2,3,4,5\\
	classifier-11 & 3 & 0,1,2,3,4,5\\
	\hline
	\end{tabular}
\end{table}

@@

\begin{table}[h]
    \caption{Accuracy of Chosen Binary-Classifiers}
    \label{tab:classifer-accuracy}
	\begin{tabular}{|c|c|c|c|c|}
	\hline
	Classifier ID & 3 & 4 & 1 & 2 \\
	\hline
	Accuracy & 96.0\% & 86.0\% & 91.5\% & 87.0\%\\
	\hline
	\end{tabular}
\end{table}

@@

\begin{table}[h]
    \caption{Encoding and Accuracy of Each Class Label}
    \label{tab:encoding-and-accuarcy}
	\begin{tabular}{|c|c|c|c|c|c|c|}
	\hline
	Class ID & 0 & 1 & 2 & 3 & 4 & Overall \\
	\hline
	Encoding &0 & 110 & 1110 &10 & 1111 &\\
	\hline
	\end{tabular}
\end{table}
